---
title: "ADMIXTOOLS-QPGRAPH"
output: 
  github_document:
    toc: true
    toc_depth: 3
date: "2025-06-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r fix-pandoc-path, include=FALSE}
Sys.setenv(PATH = paste("/opt/homebrew/bin", Sys.getenv("PATH"), sep = ":"))
Sys.setenv(RSTUDIO_PANDOC = "/opt/homebrew/bin")
```


## BACKGROUND
Admixtools is a software package for the analysis of population structure and admixture. This folder contains scripts to run QpGraph in Admixtools2, which is an R friendly version of the original Admixtools. Note that this is still different from admixr, which is another R-adapted admixtools package (although results will be practically the same).The documentation for Admixtools2 can be found [here](https://uqrmaie1.github.io/admixtools/articles/admixtools.html)

To run QpGraph, follow the scripts under Admixtools2/Scripts.
Scripts were modified from https://github.com/popgenDK/seqAfrica_wildebeest/tree/main/qpGraph.

## PREREQUISITES
### Softwares
* plink/1.90b4.9
* R with following packages installed: "admixtools", "reshape2", "plotly", "tidyverse", "igraph"

### Data
* filtered BCF file containing all populations/individuals of interest
* pruning is not strictly necessary, as stats will be calculated in block sizes. However, pruning or subsampling may drastically increase speed when working with large datasets.

## RUNNING ADMIXTOOLS/QPRAPH
### Preparing inputfiles
**0_prepare_input.sh**

This script is used to update ID's to group individuals of the same population/species and requires a *SUBSET.txt* file.

To prepare this file: 

1. copy the .fam file.
1. Modify the first column with the new ID's.

E.g. this is what my SUBSET.txt file looks like (minus the header):

| Population   | Sample ID             | Col3 | Col4 | Col5 | Col6 |
|--------------|-----------------------|------|------|------|------|
| Mouflon_IRA  | 266                   | 0    | 0    | 0    | -9   |
| Mouflon_IRA  | 267                   | 0    | 0    | 0    | -9   |
| Mouflon_IRA  | 271                   | 0    | 0    | 0    | -9   |
| Mouflon_IRA  | 272                   | 0    | 0    | 0    | -9   |
| DOM_ALT      | Altamurana-SRR12396902 | 0    | 0    | 0    | -9   |
| DOM_ALT      | Altamurana-SRR12396903 | 0    | 0    | 0    | -9   |
| Argali       | ARG2-1                | 0    | 0    | 0    | -9   |
| Argali       | ARG3-1                | 0    | 0    | 0    | -9   |
| Argali       | ARG8-2                | 0    | 0    | 0    | -9   |
| Argali       | ARG10-4               | 0    | 0    | 0    | -9   |
| Argali       | ARG19                 | 0    | 0    | 0    | -9   |
| Argali       | ARG20                 | 0    | 0    | 0    | -9   |
